---
title: "First Steps to Use the Facebook API"
author: "Sofia Gil-Clavel"
date: "Abril 2022"
output:
  md_document:
    toc: true 
    toc_depth: 3  
---

<!--  https://bookdown.org/yihui/rmarkdown/html-document.html -->

This tutorial is an adaptation of the one I gave in 2019:
https://github.com/SofiaG1l/Using_Facebook_API

**If you find the tutorial useful, then please do not forget to cite my Facebook articles:**

* Gil-Clavel, Sofia, and Emilio Zagheni. “Demographic Differentials in Facebook Usage around the World.” Proceedings of the International AAAI Conference on Web and Social Media 13 (2019): 647–50.

* Gil-Clavel, Sofia, Emilio Zagheni, and Valeria Bordone. “Close Social Networks Among Older Adults: The Online and Offline Perspectives.” Population Research and Policy Review, October 26, 2021. https://doi.org/10.1007/s11113-021-09682-3.


# First step

To generate the token we will need to enable Facebook advertising on your Facebook account.

To do this you can follow this guide, just remember you want to open a *Marketing App*: 
https://github.com/SofiaG1l/Using_Facebook_API/blob/master/First_Step.pdf

Once you have a developer account, you can use the following links to:

* Access your applications:
https://developers.facebook.com/apps/

* To monitor when your token expires: 
https://developers.facebook.com/tools/debug/accesstoken


# API Step by Step

```{r Cleaning Environment, eval=FALSE}
# Cleaning the R Environment
rm(list=ls())

# Cleaning the Computer Memory
gc()

```


## Credential Settings

```{r Setting up the credentials, eval=FALSE}

token="<Your Token>" # Change the "Your Token" with your token and delete the: <>

act="<Your Act>" # Change the "Your Act" with your Creation Act and delete the: <>

version="<vXX.X>" # Change the Xs for your version and delete the: <>

Credentials=paste0('https://graph.facebook.com/',version,'/act_',act,'/delivery_estimate?access_token=',token,'&include_headers=false&method=get&optimization_goal=REACH&pretty=0&suppress_http_code=1')

```


## The basics

### Using the Facebook Marketing Graphic User Interface

In the next pdf, I explain how to access the Graphic User Interface of the Facebook Marketing Manager. In it, I check the number of Facebook users that are German and older than 17 years old. As you can see, the Facebook Marketing Manager returns the "Estimated audience size" of Facebook users in Germany that are older than 17 years old, which in April 21st 2022 was between 40.4 million and 47.6 million.

https://github.com/SofiaG1l/Taller_COLMEX_API/blob/main/Facebook_API/FB_GUI_Audience.pdf

### 1.Using Basic URLs

First, we will try to use a browser. I would recommend you to use Mozilla, as it displays the result in a nicer manner than Google Chrome. 

What you need to do with the following link is to replace your credentials where necessary and then copy the full link into your browser. 

```{r Basic URL, eval=FALSE}

# https://graph.facebook.com/<<vXX.X>>/act_<<ACT>>/delivery_estimate?access_token=<<TOKEN>>&include_headers=false&method=get&pretty=0&suppress_http_code=1&method=get&optimization_goal=REACH&pretty=0&suppress_http_code=1&targeting_spec={"geo_locations":{"countries":["MX"]},"genders":[1,2] ,"age_min":18, "age_max":65}


```

This should return/display a type of data called JSON. In it you will see the next variables:

- daily_outcomes_curve 
- estimate_dau 
- estimate_mau_lower_bound
- estimate_mau_upper_bound 
- estimate_ready

You can learn more about them here:
https://developers.facebook.com/docs/marketing-api/audiences/reference/estimated-daily-results/

### 2.Download data programmatically

To download and transform the data to a data frame we will use the packages **httr**, **jsonlite** y **tidyverse**. If you have not installed them in R, then first you would need to use the function *install.packages("name")* replacing **name** by the name of the package.

```{r Loading Packages, eval=FALSE}

require(httr)
require(jsonlite)
require(tidyverse)

```


### 3.Total population disaggregated by age, sex and country

The parameters we use are in JSON format. You can learn more about that format in the next [linnk](https://www.w3schools.com/js/js_json_intro.asp). We will handle the format in R using strings.
  
* age_min: integer value
* age_max: integer value
* genders: integer value
* geo_locations: JSON object where *country* is a matrix

```{r Age Gender Country, eval=FALSE}

Age1=18
Age2=65

g="1,2" # 1:male and 2:female, and both is: g="1,2"

C='"MX"' # The ISO-2 code of the country

query <- paste0(Credentials,'&
                targeting_spec={
                "age_min":',Age1,',
                "age_max":',Age2,',
                "genders":[',g,'],
                "geo_locations":{"countries":[',C,'],"location_types":["home"]}}')


(query_val<-url(gsub('[\n\t ]','',query))%>%fromJSON)

t(query_val$data)

```

### 4.Total population matching certain characteristics broken down by age, sex and country

The first step is to know the name of all the possible variables that we can query. There are three different classes that you can call by changing the parameter *class*:

* demographics
* interests
* behaviors

```{r Checking FB Variables, eval=FALSE}

DF_CHARTICS<-GET(
  
  paste0("https://graph.facebook.com/",version,"/search"),
  
  query=list(
    
    type='adTargetingCategory',
    
    class='demographics',
    
    access_token=token,
    
    limit=2000
    
  )) %>%content(as="text")%>%fromJSON%>%.[[1]]

View(DF_CHARTICS)

```

Now we are going to prepare a basic query, for this you just have to choose a variable and save the following information:

```{r Basic Query with Var, eval=FALSE}

ROW=1

(TYPE=DF_CHARTICS$type[ROW])
(ID=DF_CHARTICS$id[ROW])
(NAME=DF_CHARTICS$name[ROW])

```

To segment populations that match specific characteristics we will use the *flexible_spec* parameter of the Facebook Marketing API, this parameter is a JSON object. To incorporate it into our initial string, we'll store the string in the variable **CHARTICS**.

```{r, eval=FALSE}

CHARTICS<-paste0(',"flexible_spec":[{"',TYPE,'":[{"id":"',ID,'","name":"',NAME,'"}]}]')

# A basic query including this parameter is:
  
query <- paste0(Credentials,'&
targeting_spec={"age_min":',Age1,',
                "age_max":',Age2,',
                "genders":[',g,']',
                CHARTICS,',
                "geo_locations":{"countries":[',C,'],"location_types":["home"]},
                "facebook_positions":["feed","instant_article","instream_video","marketplace"],
                "device_platforms":["mobile","desktop"],
                "publisher_platforms":["facebook","messenger"],
                "messenger_positions":["messenger_home"]}')


(query_val<-url(gsub('[\n\t ]','',query))%>%fromJSON)

t(query_val$data)

```

>Depending on the API version, you may need to delet line 200:
```{r, eval=FALSE}
"device_platforms":["mobile","desktop"],
```


In the case of specific characteristics, you can make the following types of queries:
https://developers.facebook.com/docs/marketing-api/audiences/reference/advanced-targeting#broadcategories

Here is how to translate the examples in the link to R:

* one characteristic **and**  another:

```{r AND, eval=FALSE}
'"flexible_spec":[{
                    "TYPE_1":[{"id":"ID_1","name":"NAME_1"}]
                  },
                  {
                    "TYPE_2":[{"id":"ID_2","name":"NAME_2"}]
                  }]'
```

* one feature **or** another:

```{r OR, eval=FALSE}

'"flexible_spec":[{
    "TYPE_1":[{"id":"ID_1","name":"NAME_1"}],
    "TYPE_2":[{"id":"ID_2","name":"NAME_2"}]
  }]'

```

In the case of OR we need to group by TYPE. Check the following example: 
*People who are travelers OR like soccer OR movies*

```{r OR example, eval=FALSE}

'"flexible_spec": [{ 
    "behaviors": [
          {"id":6002714895372,"name":"All travelers"}
        ], 
    "interests": [ 
          {"id":6003107902433,"name":"Association football (Soccer)"}, 
          {"id":6003139266461,"name":"Movies"} 
        ] 
  }]'

```


More info here: https://developers.facebook.com/docs/marketing-api/audiences/reference/advanced-targeting#broadcategories

```{r Full Example, eval=FALSE}

ROW=which(DF_CHARTICS$name=="Away from hometown")

(TYPE_1=DF_CHARTICS$type[ROW])
(ID_1=DF_CHARTICS$id[ROW])
(NAME_1=DF_CHARTICS$name[ROW])

ROW=which(DF_CHARTICS$name=="Friends of people with birthdays in a month")

(TYPE_2=DF_CHARTICS$type[ROW])
(ID_2=DF_CHARTICS$id[ROW])
(NAME_2=DF_CHARTICS$name[ROW])

# Preparing string of characteristics:
CHARTICS<-paste0(',"flexible_spec":[{"',TYPE_1,'":[{"id":"',ID_1,'","name":"',NAME_1,'"}]},
                 {"',TYPE_2,'":[{"id":"',ID_2,'","name":"',NAME_2,'"}]}]')

# Preparing query:
query <- paste0(Credentials,'&
targeting_spec={"age_min":',Age1,',
                "age_max":',Age2,',
                "genders":[',g,']',
                CHARTICS,',
                "geo_locations":{"countries":[',C,'],"location_types":["home"]},
                "facebook_positions":["feed","instant_article","instream_video","marketplace"],
                "device_platforms":["mobile","desktop"],
                "publisher_platforms":["facebook","messenger"],
                "messenger_positions":["messenger_home"]}')

# Retrieving:
(queryAND_val<-url(query)%>%fromJSON)


```

>Dependiendo de la versión de la API, tal vez tengas que borrar la línea: 
```{r, eval=FALSE}
"device_platforms":["mobile","desktop"],
```


You can find a more complex example in the [code](https://github.com/SofiaG1l/OlderAdultsCloseSocialNetworks) of my article publication: [Close Social Networks Among Older Adults: The Online and Offline Perspectives](https://doi.org/10.1007/s11113-021-09682-3).

## More Sofisticated Queries

### Adding variables that are not in the lists

Suppose you want to download by [educational level](https://developers.facebook.com/docs/marketing-api/audiences/reference/advanced-targeting#education_and_workplace), what you need to do is add the level in the  **CHARTICS** array as one more variable.

```{r, eval=FALSE}

EDU='"education_statuses":["1","2"]'

CHARTICS<-paste0(',"flexible_spec":[{',EDU,'},{"',TYPE,'":[{"id":"',ID,'","name":"',NAME,'"}]}]')

# A basic query including this parameter is:
  
query <- paste0(Credentials,'&
targeting_spec={"age_min":',Age1,',
                "age_max":',Age2,',
                "genders":[',g,']',
                CHARTICS,',
                "geo_locations":{"countries":[',C,'],"location_types":["home"]}}')


query_val<-url(gsub('[\n\t ]','',query))%>%fromJSON

t(query_val$data)

```

>Dependiendo de la versión de la API, tal vez tengas que borrar la línea: 
```{r, eval=FALSE}
"device_platforms":["mobile","desktop"],
```


If you need to search by school or university then you will need to download the IDS. You can download them here:

https://developers.facebook.com/docs/marketing-api/audiences/reference/targeting-search#demo

```{r Checking FB Edu Vars, eval=FALSE}

DF_CHARTICS<-GET(
  
  paste0("https://graph.facebook.com/",version,"/search"),
  
  query=list(
    
    type='adeducationschool',
    
    q='UNAM',
    
    access_token=token,
    
    limit=2000
    
  )) %>%content(as="text")%>%fromJSON%>%.[[1]]

View(DF_CHARTICS)

```

```{r, eval=FALSE}

EDU='"education_schools":[{id:125299054202386}]'

CHARTICS<-paste0(',"flexible_spec":[{',EDU,'},{"',TYPE,'":[{"id":"',ID,'","name":"',NAME,'"}]}]')

# A basic query including this parameter is:
  
query <- paste0(Credentials,'&
targeting_spec={"age_min":',Age1,',
                "age_max":',Age2,',
                "genders":[',g,']',
                CHARTICS,',
                "geo_locations":{"countries":[',C,'],"location_types":["home"]}}')


query_val<-url(query)%>%fromJSON

t(query_val$data)

```

### Geographic variables

Let's check Facebook's documentation on city search:

https://developers.facebook.com/docs/marketing-api/audiences/reference/targeting-search/#cities

As an example, we will download the data related to the cities of Mexico. In the Facebook documentation you will find this:

```{r Checking cities, eval=FALSE}

## Cities
DF_GEO<-GET(
  
  paste0("https://graph.facebook.com/",version,"/search"),
  
  query=list(
    
    location_types='city',
    
    type='adgeolocation',
    
    q='Mexico',
    
    country_code='MX', # You can delete this parameter to create an open search
    
    access_token=token,
    
    limit=100
    
  )) %>%content(as="text")%>%fromJSON%>%.[[1]]

View(DF_GEO)

```

#### From country to region

In the following link you can find the variables referring to geo-location:

https://developers.facebook.com/docs/marketing-api/audiences/reference/basic-targeting#location

Let's start with regions:

```{r Checking counties, eval=FALSE}

## Cities
DF_GEO<-GET(
  
  paste0("https://graph.facebook.com/",version,"/search"),
  
  query=list(
    
    location_types='region',
    
    type='adgeolocation',
    
    q='Mexico',
    
    country_code='MX', # You can remove this to make your query less specific
    
    access_token=token,
    
    limit=100
    
  )) %>%content(as="text")%>%fromJSON%>%.[[1]]

View(DF_GEO)

```


```{r Checking counties 2, eval=FALSE}

query <- paste0(Credentials,'&
targeting_spec={"age_min":',Age1,',
                "age_max":',Age2,',
                "genders":[',g,'],
                "geo_locations":{"regions": [{"key":"2518"}]},
                "publisher_platforms":["facebook","messenger"]}')

(GEO_QUERY<-url(gsub('[\n\t ]','',query))%>%fromJSON)


```


To download the data related to different cities you only need to include them in the matrix:

```{r Checking country, eval=FALSE}

query <- paste0(Credentials,'&
targeting_spec={"age_min":',Age1,',
                "age_max":',Age2,',
                "genders":[',g,'],
                "geo_locations":{"regions": [{"key":"2518"},{"key":"2535"}]},
                "publisher_platforms":["facebook","messenger"]}')

(GEO_QUERY<-url(gsub('[\n\t ]','',query))%>%fromJSON)


```

Now we will download data by city:

```{r Checking cities 2, eval=FALSE}

## Cities
DF_GEO<-GET(
  
  paste0("https://graph.facebook.com/",version,"/search"),
  
  query=list(
    
    location_types='city',
    
    type='adgeolocation',
    
    q='Guadalajara',
    
    country_code='MX', # You can remove this to make your query less specific
    
    access_token=token,
    
    limit=100
    
  )) %>%content(as="text")%>%fromJSON%>%.[[1]]

View(DF_GEO)

```

```{r Checking by city 3, eval=FALSE}


query <- paste0(Credentials,'&
targeting_spec={"age_min":',Age1,',
                "age_max":',Age2,',
                "genders":[',g,'],
                "geo_locations":{
                  "cities":[{"distance_unit":"mile","key":"2673660","name":"Mexico City","region":"Distrito Federal","region_id":"2513","country":"MX","radius":25}]
                },
                "publisher_platforms":["facebook","messenger"]}')

(GEO_QUERY<-url(gsub('[\n\t ]','',query))%>%fromJSON)


```


```{r Checking city 4, eval=FALSE}

query <- paste0(Credentials,'&
targeting_spec={"age_min":',Age1,',
                "age_max":',Age2,',
                "genders":[',g,'],
                "geo_locations":{"cities": [{"key":"1522110", "radius":25, "distance_unit":"mile"}]},
                "publisher_platforms":["facebook","messenger"]}')

(GEO_QUERY<-url(gsub('[\n\t ]','',query))%>%fromJSON)


```

















